Optifcal Observations of an Extraordinarily Young Type Ia Supernova iPTF16abc

### 2017-01-20 doc/ms.tex
first completed draft

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{\today}
\revised{\today}
\accepted{\today}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
\submitjournal{ApJ}

@@

  In this paper, we present observations of a young normal Type Ia supernova
  iPTF16abc. Our analysis shows that: blabla ...

@@

\ycao{TODO: include LCOGT data in both text and figures.}

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.95\textwidth]{lightcurve.pdf}
  \caption{Multi-band light curves of iPTF16abc are shown. Filters are
    denoted by different colors and observation instruments by
    different markers. The $t_{max}$ time is the B-band maximum
    determined by SALT2 (Section). The black ticks near the top of the
    figure shows epochs of spectroscopic observations.}
  \label{fig:lightcurve}
\end{figure*}

@@

After discovery, we utilized the rainbow camera of the SED
Machine (\ycao{REF}) mounted on the 60-inch telescope at Palomar
Observatory (P60) to undertake photometric observations in $g$, $r$
and $i$ filters. The image differencing against the archival SDSS
images and forced PSF photometry on the subtracted images were
performed by the Fremling Automated Pipeline
\citep{2016A&A...593A..68F}. The photometry is also calibrated to the
SDSS catalog.

Las Cambres Observatory Global Network (LCOGT) also carry out photometric
observations in the \textit{BVgri} filters with its 1-m telescope network.
\ycao{Some info about data reduction}

In space, \textit{Swift} observed iPTF16abc for 14 epochs, covering
from the very early phase to the post-peak phase. Aperture photometry
are carried out on the images taken by its Ultraviolet-Optical
Telescope (UVOT) with the usual procedures in the HEASoft and
corrected for the coincident loss and aperture loss. No pre-SN UVOT
image at the SN location is available in the \textit{Swift} archive.
Visual inspection to the UVOT images suggests that the background
galaxy light in the UVOT filters is probably negligible. No X-ray
emission was detected at the location of the SN by the X-ray Telescope
(XRT) in any of these for epochs.

The multi-color light curves of iPTF16abc are illustrated in Figure
\ref{fig:lightcurve}.  For convenience, all magnitudes are in the AB
system with a zero point of $3631$\,Jy in all filters.

Spectroscopic observations of iPTF16abc were undertaken with the
Gemini Multi-Object Spectrograph (GMOS; \citealt{2004PASP..116..425H})
on the Gemini North telescope, Low-Resolution Imaging Spectrometer
(LRIS; \citealt{1995PASP..107..375O}) on the Keck-I telescope, DEep
Imaging Multi-Object Spectrograph (DEIMOS;
\citealt{2003SPIE.4841.1657F}) on the Keck-II telescope, The Andalucia
Faint Object Spectrograph and Camera (ALFOSC\footnote{ALFOSC
  instrument webpage:
  \url{http://www.not.iac.es/instruments/alfosc/}}) on the Nordic
Optical Telescope (NOT), and X-shooter \citep{2011A&A...536A.105V} and
Ultraviolet and Visual Echelle Spectrograph (UVES;
\citealt{2000SPIE.4008..534D}) on the Very Large Telescope (VLT). The
observing log is listed in Table \ref{tab:spec_obs_log} and the
low-resolution spectral sequence is shown in Figure
\ref{fig:spec_seq}.

% currently the following table and figure do not include data from LCOGT
\begin{deluxetable*}{cccccc}
  \tablecaption{Spectroscopic observations of iPTF16abc \label{tab:spec_obs_log}}
  \tablehead{
    \colhead{Observation Date} & \colhead{SN phase} & \colhead{Telescope/Instrument} &
    \colhead{Exposure Time (s)} & \colhead{Wavelength Coverage (\AA)} & \colhead{Resolution}
  }
  \startdata
  2016 April $05.88$ & $-15.8$ & Gemini-North/GMOS &     & $3500$ -- $9500$  & $1900$ \\
  2016 April $06.51$ & $-15.1$ & Keck-II/DEIMOS & $1491$ & $5500$ -- $8100$  & $2000$ \\
  2016 April $08.51$ & $-13.1$ & Keck-II/DEIMOS & $900$  & $5500$ -- $8100$  & $2000$ \\
  2016 April $10.38$ & $-11.3$ & Keck-I/LRIS    & $300$  & $3000$ -- $10000$ & $1000$ \\
  2016 April $14.20$ & $-7.5$  & VLT/XSHOOTER   &        & $3000$ -- $25000$ & $10000$ \\
  2016 April $16.??$ & $-5.?$  & VLT/UVES       &        & covers \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D lines & $40000$ \\
  2016 April $28.??$ & $6.4$   & NOT/ALFOSC     &        & $3300$ -- $9000$  & $360$ \\
  2016 May   $10.42$ & $18.8$  & Keck-I/LRIS    & $600$  & $3000$ -- $10000$ & $1000$ \\
  2016 May   $12.03$ & $20.4$  & VLT/XSHOOTER   &        & $3000$ -- $25000$ & $10000$ \\
  \enddata
\end{deluxetable*}

\begin{figure*}[!htb]
  \centering
  \includegraphics[width=0.9\textwidth]{spec_sequence.pdf}
  \caption{Low-resolution spectra of iPTF16abc are shown in the
    chronical sequence. The phases in units of days are labeled next
    to corresponding spectra. Telluric absorption bands are grayed
    out.}
  \label{fig:spec_seq}
\end{figure*}

@@

In the highest-resolution spectrum of iPTF16abc by UVES, individual
components of both \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets show
double-absorption profiles (Figure \ref{fig:narrow_features}),
indicating two sources of absorption along the line of sight. Fitting
two Gaussian kernels to each line of the \ion{Na}{1} doublet
simultaneously leads to redshifts of $0.02313820\pm0.00000032$ and
$0.02322408\pm0.00000033$. The total equivalent widths of the
\ion{Na}{1}\,D1 and D2 lines are $0.595\pm0.009\,\textrm{\AA}$ and
$0.609\pm0.008\,\textrm{\AA}$m, respectively. Using the empirical
relation between the equivalent width of \ion{Na}{1}\,D lines and
reddening $E(B-V)$ \citep{2012MNRAS.426.1465P}, we derive
$E(B-V)=0.361\pm0.025\textrm{mag}$.

@@

@@

Extrapolation of the earliest-phase light curve with a simple model is
usually used to estimate the first light time. In theory,
\citet{1982ApJ...253..785A} derived a quadratic law for an expanding
fireball with a constant temperature. In order to account for the
temperature change, here we utilize a power-law model
\begin{equation}
  \label{eq:broken_power_law}
  f(t) \left\{
    \begin{array}{ll}
      = 0,\ \textrm{when}\ t<t_0 \\
      \propto (t-t_0)^{\alpha},\ \textrm{when}\ t>t_0
    \end{array}
  \right.\ .
\end{equation}
Since the first few detections of the SN were made in the \textit{g}
band, we perform the fit only to the \textit{g}-band light curve. We
experiment the fitting procedure with different time windows and find
that the SN flux between $t=-19\,\textrm{days}$ and
$t=-15\,\textrm{days}$ rises approximately linearly. The joint
distribution of $\alpha$ and $t_0$ for this time window is given in
the left panel of Figure \ref{fig:early_lc_fit} and the best fit model
with $\alpha = 0.94$ and $t_0=-18.47$ is shown in the right
panel. With this best-fit value of $t_0$, our first observation was
made only $0.18\,\textrm{day}$ after the first light of the SN. The
total rise time to the \textit{B}-band peak is $18.47$ days.

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.45\textwidth]{rise_time_power_law_index.pdf}
  \includegraphics[width=0.45\textwidth]{early_lc.pdf}
  \caption{Broken Power low fitting to the early $g$-band light
    curve.
    \textit{Left:} the joint distribution of $t_0$ and
    $\alpha$. The cross marker denotes the best-fit parameters. The
    inner and outer contours represent the $68\%$ and $99.7\%$
    confidence levels.
    \textit{Right:} the best-fit model of $\alpha=0.94$ and
    $t_0=-18.47\,\textrm{days}$ is illustrated against the
    data.
  }
  \label{fig:early_lc_fit}
\end{figure*}

Since $t=-15\,\textrm{days}$, the \textit{g}-band light curve rises
significantly faster than the best-fit model above, indicating a
greater value of the power-law index $\alpha$. In fact, the
light curve between $t=-14$ and $t=-8$\,days can be fitted with a
power law of index $1.40$.

A more sophisticated broken power-law model has also been used to fit
SN early-phase light curves (e.g.,
\citealt{2013ApJ...778L..15Z,2014ApJ...783L..24Z,
  2016arXiv161202097Z,2016arXiv161202097Z}). Our above result has
indicates that the \textit{g}-band light curve of iPTF16abc can also
be fitted by a broken-power law with a power index changing from
$0.94$ to $1.40$. In fact, many young SNe, including SN2011fe
\cite{2016arXiv161202097Z} and iPTF16abc, have early light curves that
follow broken power-laws \citep{2016arXiv161202725Z}). Our above
result has indicates that the \textit{g}-band light curve of iPTF16abc
can also be fitted by a broken-power law with a power index changing
from $0.94$ to $1.40$.

The initial rise of a SN depends on the distribution of $^{56}$Ni
in the ejecta. Comparing the rapid initial rise of iPTF16abc to
theoretical calculations in \citet{2016ApJ...826...96P} suggests
strong mixing of $^{56}$Ni in the iPTF16abc ejecta. The radioactive
energy from $^{56}$Ni can quickly diffuse to the SN photosphere and
power its light curve after the SN explosion.

@@

The first light of a SN is not necessarily its explotion time
$t_{exp}$.  Thanks to different deposition of $^{56}$Ni in the SN
ejecta, as pointed out in \citet{2014ApJ...784...85P}, the SN may
experience a dark period before the radioactive energy from the
shallowest $^{56}$Ni layer escapes the ejecta. Therefore, 
\citet{2014ApJ...784...85P} suggests to examine evolution of absorption
line velocities, because the velocities do not depend on the $^{56}$Ni
mixing and are expected to evolve roughly as $v\propto(t-t_{exp})^{-0.22}$.

We focus on measuring velocities of the strongest \ion{Si}{2}\,6355
line, as the second most commonly used \ion{Ca}{2} IR triplet is very
weak in iPTF16abc. Visual inspection of the early-phase spectra shows
no sign of any high-velocity component of the \ion{Si}{2} line and
that the red wing of \ion{Si}{2} overlaps the \ion{C}{2}\,6580
line. Therefore, we fit simultaneously two gaussian kernels for the
two lines and a linear term to account for the continuum component.
Then the velocities are calculated at the minimum of the \ion{Si}{2}
gaussian kernel.

\begin{figure*}[!thb]
  \centering
  \includegraphics[width=0.45\textwidth]{exp_date_chi2.pdf}
  \includegraphics[width=0.45\textwidth]{SiIIVelocity.pdf}
  \caption{Constraints on $t_{exp}$ from fitting the velocity
    evolution of \ion{Si}{2}. \textit{Left panel:} the dashed, solid
    and dash-dotted curves show $\chi^2$ for fitting power laws with
    indices $-0.20$, $-0.22$ and $-0.24$, respectively. The red
    vertical line and the orange region indicate $t_0$ and its
    3-$\sigma$ confidence interval from Section
    \ref{sec:lc_fit}. \textit{Right panel:} Observed \ion{Si}{2}\,6355
    velocities with the best-fit power-law velocity with an index of
    $-0.22$.}
  \label{fig:velocity_t_exp}
\end{figure*}

Fitting the measured velocities of \ion{Si}{2}\,6355 line to the
$v\propto(t-t_{exp})^{-0.22}$ model leads to the best-fit explosion
time $t_{exp}=-17.95\,\textrm{days}$ with a 3-$\sigma$ confidence
interval between $-17.4\,\textrm{days}$ and $-18.3\,\textrm{days}$
(Figure \ref{fig:velocity_t_exp}). We further test the robustness of
a fixed power-law index by altering the index to $-0.20$ and $-0.24$
and find consistent results within respective 3-$\sigma$ confidence
intervals.

The derived $t_{exp}$ is roughly consistent to $t_0$ estimated in
Section \ref{sec:lc_fit}, confirming that the dark period of iPTF16abc
is very short. This conclusion is consistent with the inference from the
rapid initial rise of iPTF16abc in Section \ref{sec:lc_fit}.

Among the several SNe with measurements of $t_{exp}$ and $t_0$
\citep{2014ApJ...784...85P,2016ApJ...826..144S}, iPTF16abc has the
shortest dark period and longest rise time ($18.47$ days). In
comparison, SN2011fe is inferred to have a dark period of $\sim 1$
day \citep{2014ApJ...784...85P} and a rise time of $17.7$ days
\citep{2013A&A...554A..27P}; ASASSN-14lp has a dark period of a couple
of days and a rise time of $16.94$ days. This seems to suggest that
the time difference between SN explosion and the \textit{B}-band maximum,
which equals to summation of the dark period and the rise time, is a
constant for a SN Ia. This is perhaps because the \textit{B}-band maximum
time only depends on the total amount of $^{56}$Ni.

@@

Interestingly, the first few spectra of iPTF16abc show clear detection
of \ion{C}{2}\,6580 and \ion{C}{2}\,7234 lines. Their velocity
evolution is shown in Figure \ref{fig:velocity_t_exp} as well.

The \ion{C}{2} lines only appear in the very early phases. The
equivalent widths of detected \ion{C}{2}\,6580 and \ion{C}{2}\,7234
lines are shown in Figure \ref{fig:ew}, and are compared to that of
the \ion{Si}{2}\,6355 line. Especially in the first spectrum taken at
$t=-15.8$ days, the \ion{C}{2} equivalent widths are stronger than
that of \ion{Si}{2}.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.45\textwidth]{EW.pdf}
  \caption{Equivalent Width of \ion{Si}{2}\,6355, \ion{C}{2}\,6580 and
    \ion{C}{2}\,7234 lines}
  \label{fig:ew}
\end{figure}

The strong and short-lived \ion{C}{2} features suggest that
the outermost layers of the ejecta contain a certain amount of
carbon which is responsible for the \ion{C}{2} absorption. The
abundance of carbon drops dramatically into the inner layers
of the ejecta.

Carbon signatures are seen in over $1/4$ of all normal SNe Ia before
maxima
\citep{2011ApJ...732...30P,2012MNRAS.425.1917S,2011ApJ...743...27T},
but the signatures are usually weak. Even in SN2011fe, the \ion{C}{2}
features are not strong in the first spectra
\citep{2012ApJ...752L..26P}.  The only normal SN Ia known to have
strong \ion{C}{2} features at very early phases is SN2013dy
\citep{2013ApJ...778L..15Z}. Unlike iPTF16abc, the equivalent widths
of \ion{C}{2} features in SN2013dy are as large as that of
\ion{Si}{2}\,6355.

\subsection{Discussions and Physical Constraints}
\label{sec:lc_energy}

The early radiation of a SN Ia may have multiple resources: SN
shockbreakout, SN-companion collision, and radioactive activity. Since
each provides interesting constraints on the progenitor properties, we
explore all the three posibilities.

@@

The shock breakout of a SN Ia lasts for a fraction of a second due to
the small size of the exploding star. However, the subsequent cooling
phase may last longer (e.g., \citealt{2010ApJ...708..598P}).
Following the analysis of SN2011fe in \citet{2012ApJ...744L..17B}, we
compare the early-phase \textit{g}-band light curve of iPTF16abc with
two cooling models \citep{2011ApJ...728...63R, 2010ApJ...708..598P}
and reach a not very constraining conclusion that the radius of the
progenitor star of iPTF16abc should be $<1\sr$. In fact, the cooling
emission is negligible at the time of the first detection of
iPTF16abc.

\subsubsection{SN companion}

>>>>

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{\today}
\revised{}
\accepted{}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
\submitjournal{ApJ}

@@

  We report discovery of an extraordinarily young Type Ia supernova
  iPTF16abc by the intermediate Palomar Transient Factory. Our first
  observation was made only $0.18$ days after the onset of the
  supernova explosion. Our analysis shows that the early emission of
  this supernova dominantly comes from the radioactive decay instead
  of supernova shock breakout or collision between the ejecta and a
  possible companion star. The fast initial rise of the supernova, the
  velocity evolution of spectral lines, and the strong and short-lived
  carbon feature in the earliest spectra together provide strong
  evidence of strong mixing of synthesized radioactive elements during
  the supernova explosion.

@@

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.95\textwidth]{lightcurve.pdf}
  \caption{Multi-band light curves of iPTF16abc are shown. Filters are
    denoted by different colors and observation instruments by
    different markers. The $t_{max}$ time is the B-band maximum
    determined by SALT2 (Section \ref{sec:classification}). The solid
    curves are best-fit results from SALT2. The black ticks near the
    top of the figure shows epochs of spectroscopic observations.}
  \label{fig:lightcurve}
\end{figure*}

@@

After discovery, we utilized the rainbow camera of the SED Machine
(\ycao{REF}) mounted on the 60-inch telescope at Palomar Observatory
(P60) to undertake photometric observations in $g$, $r$ and $i$
filters. The image differencing against the archival SDSS images and
forced PSF photometry on the subtracted images were performed by the
Fremling Automated Pipeline \citep{2016A&A...593A..68F}. The
photometry is also calibrated to the SDSS catalog \ycao{which DR?}.

Las Cambres Observatory Global Network (LCOGT) also carry out
photometric observations in the \textit{BVgri} filters with its 1-m
telescope network.  PSF photometry was measured on these images using
the \texttt{lcogtsnpipe} pipeline \citep{2016MNRAS.459.3939V}. The
\textit{BV} magnitudes are calibrated to the Fourth USNO CCD
Astrograph Catalog \citep{2013AJ....145...44Z}, and the \textit{gri}
magnitudes are calibrated to SDSS Data Release 6
\citep{2008ApJS..175..297A}.

In space, \textit{Swift} observed iPTF16abc for 14 epochs, covering
from the very early phase to the post-peak phase. Aperture photometry
are carried out on the images taken by its Ultraviolet-Optical
Telescope (UVOT) with the usual procedures in the HEASoft and
corrected for the coincident loss and aperture loss. The image counts
are converted to physical fluxes using the latest calibration
\citep{2011AIPC.1358..373B}. No pre-SN UVOT image at the SN location
is available in the \textit{Swift} archive.  Visual inspection to the
UVOT images suggests that the background galaxy light in the UVOT
filters is probably negligible. No X-ray emission was detected at the
location of the SN by the X-ray Telescope (XRT) in any of these
epochs.

The multi-color light curves of iPTF16abc are illustrated in Figure
\ref{fig:lightcurve}.  For convenience, magnitudes in \textit{all}
filters are in the AB system with a zero point of $3631$\,Jy.

Spectroscopic observations of iPTF16abc were undertaken with a variety
of telescopes and instruments in multiple epochs spanning from a
couple of days after explosion to two months after its maximum. An
observing log is listed in Table \ref{tab:spec_obs_log}. Data
reduction is undertaken with usual routines in IDL/Python. Except for
the high-resolution spectra obtained by VLT instruments, the optical
spectral evolution of iPTF16abc is illustrated in Figure
\ref{fig:spec_seq}.

\begin{deluxetable*}{cccccc}
  \tablecaption{Spectroscopic observations of iPTF16abc \label{tab:spec_obs_log}}
  \tablehead{
    \colhead{Observation MJD} & \colhead{SN phase} & \colhead{Telescope} &
    \colhead{Instrument} & \colhead{Wavelength Coverage (\AA)}
  }
  \startdata
  $57483.26$ & $-16.4$ & DCT & DeVeny\tablenotemark{1} & $3301$--$7499$ \\
  $57483.88$ & $-15.8$ & Gemini-North & GMOS\tablenotemark{2} & $3800$--$9200$ \\
  $57484.51$ & $-15.1$ & Keck-II & DEIMOS\tablenotemark{3} & $5500$--$8099$ \\
  $57486.51$ & $-13.1$ & Keck-II & DEIMOS\tablenotemark{3} & $5500$--$8099$ \\
  $57488.38$ & $-11.3$ & Keck-I & LRIS\tablenotemark{4} & $3055$--$10411$ \\
  $57489.51$ & $-10.1$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3301$--$8999$ \\
  $57490.40$ & $ -9.3$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3301$--$9999$ \\
  $57491.55$ & $ -8.1$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3300$--$9998$ \\
  $57492.20$ & $ -7.5$ & VLT & X-shooter\tablenotemark{6} & $3300$--$24550$ \\
  $57494.00$ & $ -5.7$ & VLT & UVES\tablenotemark{7} & \\
  $57503.32$ & $ +3.7$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3300$--$9999$ \\
  $57506.00$ & $ +6.3$ & NOT & ALFOSC\tablenotemark{8} & $3602$--$8098$ \\
  $57508.27$ & $ +8.6$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $3301$--$9999$ \\
  $57518.42$ & $+18.8$ & Keck-I & LRIS\tablenotemark{4} & $3071$--$10208$ \\
  $57520.03$ & $+20.4$ & VLT & X-shooter\tablenotemark{6} & $3300$--$24789$ \\
  $57529.40$ & $+29.8$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $4000$--$8998$ \\
  $57542.41$ & $+42.8$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $4000$--$8998$ \\
  $57550.40$ & $+50.8$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $4001$--$8999$ \\
  $57562.38$ & $+62.7$ & LCOGT-2m & FLOYDS\tablenotemark{5} & $4800$--$9300$ \\
  \enddata
  \tablenotetext{1}{The Deveny Spectrograph \citep{2014SPIE.9147E..2NB}}
  \tablenotetext{2}{The Gemini Multi-Object Sectrograph \citep{2004PASP..116..425H}}
  \tablenotetext{3}{DEep Imaging Multi-Object Spectrograph \citep{2003SPIE.4841.1657F}}
  \tablenotetext{4}{Low-Resolution Imaging Spectrometer \citep{1995PASP..107..375O}}
  \tablenotetext{5}{FLOYDS \url{https://lco.global/observatory/instruments/floyds}}
  \tablenotetext{6}{X-shooter \citep{XShooter}}
  \tablenotetext{7}{Ultraviolet and Visual Echelle Spectrograph \citep{2000SPIE.4008..534D}}
  \tablenotetext{8}{The Andalucia Faint Object Spectrograph and Camera \url{http://www.not.iac.es/instruments/alfosc}}
\end{deluxetable*}

\begin{figure*}[!htb]
  \centering
  \includegraphics[width=1.0\textwidth]{spectra.pdf}
  \caption{Low-resolution spectra of iPTF16abc are shown in the
    chronical orders. In order for better illustration, each spectrum
    is normalized by the median flux value between $6,000$ and
    $7,000\,\textrm{\AA}$ and offset properly.  The phases in units of
    days are noted next to corresponding spectra. Telluric absorption
    bands are grayed out. The narrow \ion{Na}{1}\,D absorption is also
    highlighted in orange.}
  \label{fig:spec_seq}
\end{figure*}

@@

In the highest-resolution spectrum of iPTF16abc by UVES, individual
components of both \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets show
double-absorption profiles (Figure \ref{fig:narrow_features}),
indicating two sources of absorption along the line of sight. Fitting
two Gaussian kernels to each line of the \ion{Na}{1} doublet
simultaneously leads to redshifts of $0.02313820\pm0.00000032$ and
$0.02322408\pm0.00000033$. The total equivalent widths of the
\ion{Na}{1}\,D1 and D2 lines are $0.595\pm0.009\,\textrm{\AA}$ and
$0.609\pm0.008\,\textrm{\AA}$m, respectively. Using the empirical
relation between the equivalent width of \ion{Na}{1}\,D lines and
reddening $E(B-V)$ \citep{2012MNRAS.426.1465P}, we derive
$E(B-V)=0.361\pm0.025\textrm{mag}$. We caution that the empirical
relation has a quite large scatter.

@@

We also note in Figure \ref{fig:narrow_features} that the absorption
profiles of \ion{Ca}{2}\,H$+$K are quite different from those of \ion{Na}{1}\,D
doublet, implying that the dust compositions in the two absorption sources
may have some difference.

@@

Extrapolation of the earliest-phase light curve with a simple
mathematical model is usually used to estimate the first light time of
a SN. In theory, \citet{1982ApJ...253..785A} derived a quadratic law
for an ideal expanding fireball with a constant temperature. In order
to account for variation of photospheric temperatures during
expansion, we utilize a power-law curve to model the early light curve.
\begin{equation}
  \label{eq:broken_power_law}
  f(t) \left\{
    \begin{array}{ll}
      = 0,\ \textrm{when}\ t<t_0 \\
      \propto (t-t_0)^{\alpha},\ \textrm{when}\ t>t_0
    \end{array}
  \right.\ .
\end{equation}
Since the first few detections of the SN were made in the \textit{g}
band, our modeling is limited to the early \textit{g}-band light
curve. We minimize $\chi^2$ in a large search grid of $t_0$, $\alpha$
and the proportionality constant. We also note that the definition of
the time window for the early light curve may affect the fitting
result, so we experiment the fitting procedure with different time
windows. The modeling results show that the SN flux between
$t=-19\,\textrm{days}$ and $t=-15\,\textrm{days}$ rises approximately
linearly. Figure \ref{fig:early_lc_fit} shows the best-fit result and
the joint marginal distribution of $t_0$ and $\alpha$. With the
best-fit model of $\alpha=0.92$ and $t_0=-18.47$ days, our first
observation was made only $0.18\,\textrm{day}$ after the first light
of the SN. The total rise time to the \textit{B}-band peak is $18.47$
days.

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.95\textwidth]{early_lc.pdf}
  \caption{Broken Power low fitting to the early $g$-band light
    curve. The best-fit model of $\alpha=0.92$ and $t_0=-18.47\,\textrm{days}$
    and corresponding residues are shown in the top and bottom panels, respectively.
    The joint distribution of $t_0$ and $\alpha$ is illustrated in the inset of
    the upper panel. The solid and dashed contours represent the $68\%$ and $99.7\%$
    confidence levels.
  }
  \label{fig:early_lc_fit}
\end{figure*}

Starting from $t=-15\,\textrm{days}$, the \textit{g}-band light curve
rises significantly faster than the best-fit model above, indicating a
greater value of the power-law index $\alpha$. In fact, the light
curve between $t=-14$ and $t=-8$\,days can be fitted with another
power law of index $1.40$. In another word, the whole light curve
before $t=-8$\,days can be approximated by a broken power-law model
(e.g.,
\citealt{2013ApJ...778L..15Z,2014ApJ...783L..24Z,2016arXiv161202097Z,
  2016arXiv161202725Z}).

As discussed in Section \ref{sec:lc_energy}, the initial light curve
of iPTF16abc is probably purely powered by radioactive decay of
$^{56}$Ni. Therefore, the initial rise of the light curve depends on
the depth of the shallowest layer where $^{56}$Ni is deposited
\citep{2014ApJ...784...85P}. According to theoretical mdels
\citep{2016ApJ...826...96P}, the linearly rising light curve implies
strong mixing of $^{56}$Ni in the ejecta. As a result, the time lag
between the SN explosion and the initial rise of the radioactively
powered light curve is negligible.

@@

In order to constrain the explosion time of the SN independently of
the SN light curve, \citet{2014ApJ...784...85P} suggested to measure
the expansion velocities of the photosphere, because the expansion of
the ejecta starts at the onset of the SN explosion. Assuming a
constant opacity in the ejecta, the photospheric velocity evolves as
$v_{ph}\propto(t-t_{exp})^{-0.22}$. While the photospheric velocities
are not easy to measure, line velocities of \ion{Si}{2} or \ion{Ca}{2}
are usually used to approximate the photospheric velocities
\citep{2014ApJ...784...85P,2016ApJ...826..144S}.

In the case of iPTF16abc, the \ion{Ca}{2} IR triplet is very weak in
the spectra probably due to high temperatures in the ejecta. Thus we
perform our measurements on the \ion{Si}{2}\,6355 line. Visual
inspection of the spectra shows no sign of multi-velocity components
of \ion{Si}{2}, and that the \ion{C}{2}\,6580 line overlaps the red
wing of the \ion{Si}{2} line. Consequently we apply a mathematical
model of two gaussian kernels superposed on a linear term, which
accounts for the absorption of \ion{Si}{2} and \ion{C}{2} and the
continuum component, respectively, to approach the observed spectra
between $5,900$ and $6,500$\,\AA (rest-frame). The expansion
velocity of \ion{Si}{2} is calculated at the central wavelength
of the \ion{Si}{2} Gaussian kernel.

\begin{figure*}[!thb]
  \centering
  \includegraphics[width=0.45\textwidth]{Chi2.pdf}
  \includegraphics[width=0.45\textwidth]{VelocityPlot.pdf}
  \caption{Constraints on $t_{exp}$ from fitting the velocity
    evolution of \ion{Si}{2}.
    \textit{Left panel:} the dashed, solid
    and dash-dotted curves show $\chi^2$ for fitting power laws with
    indices $-0.20$, $-0.22$ and $-0.24$, respectively. The red
    vertical line and the orange region indicate $t_0$ and its
    3-$\sigma$ confidence interval from Section
    \ref{sec:lc_fit}.
    \textit{Right panel:} Observed \ion{Si}{2}\,6355
    velocities and the best-fit power-law velocity with an index of
    $-0.22$.}
  \label{fig:velocity_t_exp}
\end{figure*}

We fit the measured velocities of \ion{Si}{2}\,6355 line to the
$v\propto(t-t_{exp})^{-0.22}$ model by minimizing the $\chi^2$ value
and find the best-fit explosion time $t_{exp}=-17.95\,\textrm{days}$
with a 3-$\sigma$ confidence interval between $-17.4\,\textrm{days}$
and $-18.3\,\textrm{days}$ (Figure \ref{fig:velocity_t_exp}). We also
alter the power-law index to $-0.20$ and $-0.24$ to examine how
sensitively the result depends on the assumed power-law index.  we
find consistent results within respective 3-$\sigma$ confidence
intervals.

Keeping in mind that $t_{exp}$ has a very large uncertainty mostly due
to the assumption in the $v\propto(t-t_{exp})^{-0.22}$ model, we
compare the estimated $t_{exp}$ to $t_0$ (left panel of Figure
\ref{fig:velocity_t_exp}) and find that $t_0\lessapprox t_{exp}$.
Since physical causality requires $t_{exp}<t_0$, we draw a qualitative
conclusion that $t_0\simeq t_{exp}$. This conclusion is consistent
with our inference from the early light curve analysis in Section
\ref{sec:lc_fit}.

With $t_{exp}$, we estimate the actual rise time of iPTF16abc from the
SN explosion to its \textit{B}-band maximum to be $17.95$ days.  In
comparison, SN2011fe has $t_0=-17.7$ days \citep{2013A&A...554A..27P}
with a preceding dark period of $\sim 1$ day
\citep{2014ApJ...784...85P}. ASASSN-14lp has $t_0=-16.94$ days with a
preceding dark period of about acouple of days
\citep{2016ApJ...826..144S}. This comparison suggests that the actual
rise time of a SN Ia between the SN explosion and the \textit{B}-band
maximum is roughly a constant, which largely depends on the total mass
of synthesized $^{56}$Ni.

@@

Another intriguing spectral feature of iPTF16abc is strong absorption of
\ion{C}{2}\,6580 and \ion{C}{2}\,7234 lines. Our analysis on the
\ion{Si}{2}\,6355 line also allows us to measure the velocities and
pseudo-equivalent widths (pEWs) of \ion{C}{2}\,6580. We also fit a Gaussian
kernel superposing on a linear term to the spectral region near
the \ion{C}{2}\,7234 line to measure its velocity and pseudo-equivalent
width. The velocity evolution of \ion{C}{2} is shown in the right panel
of Figure \ref{fig:velocity_t_exp} and the pEW evolution
in Figure \ref{fig:ew}.

Our measurements show that the strong \ion{C}{2} absorption
only appears in the very early phases. In the spectrum taken at
$t=-15.8$ days, the pEW of \ion{C}{2}\,6580 is comparable to that
of \ion{Si}{2}\,6355. Then the pEW of \ion{C}{2} decreases
until they disappear at about $t=-10$ days.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.45\textwidth]{pEW.pdf}
  \caption{Pseudo-equivalent width of \ion{Si}{2}\,6355,
    \ion{C}{2}\,6580 and \ion{C}{2}\,7234 lines. The uncertainties
    in these measurements are less than the sizes of these symbols.}
  \label{fig:ew}
\end{figure}

Formation of \ion{C}{2} absorption requires existence of both unburned
carbon and $^{56}$Ni, the latter of which is responsible to ionize the
carbon atoms. The strong and short-lived \ion{C}{2} features observed
in the spectra of iPTF16abc show that a certain amount of carbon atoms
exist in the outer layers of the ejecta and that $^{56}$Ni atoms are also
strongly mixed in the same layers. This provides a third piece of evidence
to strong mixing in the iPTF16abc ejecta.

Carbon signatures are seen in over $1/4$ of all normal SNe Ia before
maxima
\citep{2011ApJ...732...30P,2012MNRAS.425.1917S,2011ApJ...743...27T},
but the signatures are usually weak. Even in SN2011fe, the \ion{C}{2}
features are not strong in the first spectra
\citep{2012ApJ...752L..26P}.  The only normal SN Ia known to have
strong \ion{C}{2} features at early phases is SN2013dy
\citep{2013ApJ...778L..15Z}. Unlike iPTF16abc, however, the equivalent
widths of \ion{C}{2} features in SN2013dy are as large as those of
\ion{Si}{2}\,6355 in the same spectra.

\subsection{Discussions}
\label{sec:lc_energy}

Early radiation of a SN Ia may have multiple energy sources: SN shock
breakout, SN-companion collision, and radioactive activity. The
previous analysis is based on the assumption that the light curve of
iPTF16abc is powered purely by radioactive decay. Here we discuss the
other two possibilities.

@@

The shock breakout of a SN Ia lasts for a fraction of a second due to
the small size of the exploding star. However, the subsequent cooling
phase may last longer (e.g., \citealt{2010ApJ...708..598P}).
Following the analysis of SN2011fe in \citet{2012ApJ...744L..17B}, we
compare the early-phase \textit{g}-band light curve of iPTF16abc with
two cooling models \citep{2011ApJ...728...63R, 2010ApJ...708..598P}
and reach a not very constraining conclusion that the radius of the
progenitor star of iPTF16abc should be $<1\sr$. In fact, with a
typical $0.01\sr$ radius of a white dwarf, the cooling emission of a
SN Ia is negligible at the time of the first detection of iPTF16abc.

\subsubsection{SN-Companion collision}

From a SN Ia born in a single-degenerate system, with an odd of
$\lesssim10\%$ due to the geometry, we expect to see emission produced
by SN ejecta slamming into the companion. According to calculations by
\citet{2010ApJ...708.1025K}, collision between SN ejecta and the
companion star generates thermal emission with a spectrum with
spectrum that peaks in the ultraviolet. The Rayleigh-Jeans tail
emission in the \textit{g} band is very weak.
